---
title: "Color-issues in hydrological publications"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(pillar.min_chars = 15)
```

## Load data frame

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r read_data}
#read data remotely from github
df <- read_tsv("https://raw.githubusercontent.com/modche/rainbow_hydrology/main/hess_papers_rainbow.txt")
```

## 1. Overwiew data variables

-   year = year of publication (YYYY)
-   date = date (YYYY-MM-DD) of publication
-   title = full paper title from journal website
-   authors = list of authors comma-separated
-   n_authors = number of authors (integer between 1 and 27)
-   col_code = color-issue classification (see below)
-   volume = Journal volume
-   start_page = first page of paper (consecutive)
-   end_page = last page of paper (consecutive)
-   base_url = base url to access the PDF of the paper with `/volume/start_page/year/`
-   filename = specific file name of the paper PDF (e.g. `hess-9-111-2005.pdf`)

```{r structure}
str(df)
head(df)
tail(df)
skimr::skim(df)
```

### 2. Explanatory data analysis

Distribution of surveyed papers and 2005, 2010, 2015 and 2020.

```{r}
df %>% count(year)
```

Color classification is stored in the `col_code` variable with:

-   `0` = chromatic and issue-free,
-   `1` = red-green issues,
-   `2`= rainbow issues and
-   `bw`= black and white paper.

```{r col_code}
df %>% count(col_code)
```

Figure showing number of authors across color classification.

```{r col_code_fig, fig.height = 2.5, echo = FALSE}
ggplot(data = df) +
    geom_histogram(aes(x = n_authors), 
                   fill = "grey30", 
                   binwidth = 1, 
                   colour = "white",
                   size=.5) +
    facet_wrap(~col_code, nrow = 1) +
    theme_bw(14) +
    theme(panel.grid = element_blank())
    
```

### 3. Access PDF papers with download links

Data frame can be accessed to extract a vector of links to download specific papers.

**Example:** Access a specific paper from 2005:

```{r}
df %>% filter(year == 2005, start_page == 111) %>% 
    select(base_url, filename) %>% 
    mutate(download_link = paste0(base_url, filename)) %>% 
    pull(download_link)
```

**Example:** Single-author papers from 2005 that are pure black and white papers:

```{r}
df %>% filter(year == 2005, col_code == "bw", n_authors == 1) %>% 
    select(base_url, filename) %>% 
    mutate(download_link = paste0(base_url, filename)) %>% 
    pull(download_link)
```

**Example**: Rainbow papers from 2020 with more than 10 authors:

```{r}
df %>% filter(year == 2020, col_code == 2, n_authors > 10) %>% 
    select(base_url, filename) %>% 
    mutate(download_link = paste0(base_url, filename)) %>% 
    pull(download_link)
```

### 4. Other possibilities for data analysis:

```{r}
df %>% filter(str_detect(string = authors, pattern = "Weiler"))
df %>% filter(str_detect(string = title, pattern = "radar"))
df %>% filter(end_page > start_page + 30)
```
